Searching for long-period transiting planets in the \kepler\ light curves using supervised classification

### 2015-04-07 document/ms.tex
adding prediction plot and some words about expected number of single transits

% Notation for this paper.
\newcommand{\flux}{{\ensuremath{f}}}
\newcommand{\ferr}{{\ensuremath{\sigma_\flux}}}
\newcommand{\attime}{{\ensuremath{t}}}
\newcommand{\basis}{{\bvec{A}}}
\newcommand{\weights}{{\bvec{w}}}

\newcommand{\period}{{\ensuremath{P}}}
\newcommand{\phase}{{\ensuremath{T^0}}}
\newcommand{\duration}{{\ensuremath{D}}}
\newcommand{\depth}{{\ensuremath{Z}}}
\newcommand{\transittime}{{\ensuremath{T}}}
\newcommand{\impact}{{\ensuremath{b}}}
\newcommand{\ecc}{{\ensuremath{e}}}
\newcommand{\pomega}{{\ensuremath{\omega}}}

@@

@@

The \kepler\ Mission measured photometry for about 150,000 stars at half-hour
cadence for a baseline of over four years.
We aim to search these time series for single transits of long-period planets
and single eclipses of binary stars.
These data are made available on
MAST\footnote{\url{https://archive.stsci.edu/kepler/}} and, for each target,
we downloaded the full set of light curve files provided by Data Release 24
\citep{Thompson:2015}.
From these files, we extracted the PDC time series and split them into
``sections'' with no more than two contiguous missing or flagged data points.
The PDC light curves have been corrected for the instrumental effects caused
by the spacecraft using a data-driven model of the focal plane
\citep{Stumpe:2012, Smith:2012}.
Crucially, an attempt is also made by the PDC procedure to remove sharp
instrumental artifacts like ``sudden pixel sensitivity dropouts (SPSDs)''.
Any remaining missing data points are filled in using linear interpolation
resulting in sections of contiguous time series measurements with nearly
uniform time sampling.
This procedure yields about 150 \emph{light curve sections} for the typical
\kepler\ target.

@@

@@

>>>>

% Notation for this paper.
\newcommand{\period}{{\ensuremath{P}}}
\newcommand{\rp}{{\ensuremath{R_\mathrm{P}}}}
\newcommand{\rate}{{\ensuremath{\Gamma}}}

@@

\section{Estimated yield}

Before the launch of the \kepler\ Mission, \citet{Yee:2008} predicted the
yield of single transit events based on the planet occurrence rates estimated
based on small catalog of radial velocity discoveries \citep{Butler:2006}.
Using these early results and the pre-launch specifications of the \kepler\
Mission, \citet{Yee:2008} predicted that \kepler\ would discover $\sim 6$
single transit events in the full dataset.
Now that the Mission is complete and we have better estimates of the
occurrence rate and distribution of planets \citep[for example][]{Dong:2013,
Petigura:2013, Foreman-Mackey:2014, Dressing:2015}, we can update the
predicted yield for single transit events in the \kepler\ light curves.
To make this estimate, we extrapolate the distribution of large planets on
relatively short orbits out to longer periods, taking detection efficiency
and the survey and targeting properties into account.

We will base our estimate on an assumed model $Q_k(\rp,\,\period)$ for the
absolute probability of detecting a planet with radius \rp\ and period
\period\ orbiting the star $k$, and the occurrence rate distribution
\begin{eqnarray}
\rate (\rp,\,\period) &=& \frac{\dd N}{\dd\ln\rp\dd\ln\period} \quad,
\end{eqnarray}
the expected number of planets per star, per logarithmic radius, per
logarithmic period.
Given these two quantities, the expected number of single transits in the
\kepler\ data is given by
\begin{eqnarray}
N &=& \sum_{k=1}^K N_k
\end{eqnarray}
where the sum is over the $K$ stars in the sample and $N_k$ is the expected
number of observable transits around star $k$
\begin{eqnarray}\eqlabel{nk}
N_k &=& \int Q_k(\rp,\,\period) \, \rate (\rp,\,\period)
    \dd\ln\rp \dd\ln\period \quad.
\end{eqnarray}
The integral in \eq{nk} is over the full range of target parameters.

For the purposes of this discussion, we assume an approximate simple detection
efficiency model with three contributions: the geometric transit probability,
the temporal transit probability, and signal-to-noise ratio threshold of the
search technique.
Assuming circular orbits, the geometric transit probability is given by
\citep{Winn:2010}
\begin{eqnarray}
Q_k^\mathrm{(geom)}(\rp,\,\period) &=& \frac{R_k}{a} \\
&=& \left( \frac{4\,\pi^2}{G\,M_k} \right)^{1/3} \, R_k \, \period^{-2/3}
\end{eqnarray}
where $R_k$ and $M_k$ are the radius and mass of the star $k$ respectively.
The eccentricity distribution of these long-period planets will affect this
transit probability \citep{Kipping:2014} but this simple prescription should
be sufficient for a rough estimate.
For long-period orbits, the temporal transit probability will be given by
\begin{eqnarray}
Q_k^\mathrm{(time)}(\rp,\,\period) &=& \frac{T_k}{\period}
\end{eqnarray}
where $T_k$ is the total time that \kepler\ spent observing the star $k$.
Finally, the detection threshold depends on the detailed sensitivity of the
search procedure but we will approximate it as a simple step function in
signal-to-noise ratio of the transit.
This contribution will be given approximately by
\begin{eqnarray}
Q_k^\mathrm{(detect)}(\rp,\,\period) &=& \left\{\begin{array}{ll}
1 & \mathrm{if}\,\left(\rp/R_k\right)^2 > f\,\sigma_k \\
0 & \mathrm{otherwise}
\end{array}\right.
\end{eqnarray}
where $\sigma_k$ is an estimate of the noise in light curve of star $k$ and
$f$ is the detection threshold for the method.

Combining the detection efficiency components, the integral from \eq{nk}
becomes
\begin{eqnarray}
N_k &=& \left( \frac{4\,\pi^2}{G\,M_k} \right)^{1/3} \, R_k \, T_k
    \int_{\rp_\mathrm{min}} ^{\rp_\mathrm{max}} \frac{\dd\rp}{\rp}
    \int_{\period_\mathrm{min}} ^{\period_\mathrm{max}}
        \period^{-8/3}\,\rate(\rp,\,\period) \dd\period
\end{eqnarray}
where all the integration limits are set by the target parameter space.
Because of the detection probability threshold, $\rp_\mathrm{min}$ can be no
smaller than $R_k\,\sqrt{f\,\sigma_k}$.
\citet{Dong:2013} used the catalog of short period transiting planets found
by \kepler\ to constrain a model for the occurrence rate of large planets of
the form
\begin{eqnarray}
\rate(\rp,\,\period) &=& C\,\left(\frac{\period}{10\,\mathrm{d}}\right)^\beta
\end{eqnarray}
in a set of radius bins.

Assuming the stellar parameters provided by the NASA Exoplanet
Archive\footnote{\dfmtodo{URL for kepler stellar table; downloaded on SOME
DATE.}} \dfmtodo{cite Huber} and approximating the stellar noise using the
15-hour CDPP \dfmtodo{cite CDPP}, \dfmtodo{some figure} shows the extrapolated
number of transiting planets with $\period > 1500\,\mathrm{d}$ based on the
\citet{Dong:2013} power-law model.
For a detection threshold $f\sim10$, the expected number of single transit
events in the \kepler\ light curves is $\sim 120$.

\begin{figure}[p]
\begin{center}
\includegraphics{figures/predict.pdf}
\end{center}
\caption{%
The expected number of single transit events---extrapolated from the
\citet{Dong:2013} power-law model fit to the shorter period \kepler\
candidates---as a function of the effective signal-to-noise threshold of the
search procedure.
The shaded region indicates the uncertainties propagated from the model
parameters.
\figlabel{predict}}
\end{figure}

@@

The \kepler\ Mission measured photometry for about 150,000 stars at half-hour
cadence for a baseline of over four years.
We aim to search these time series for single transits of long-period planets
and single eclipses of binary stars.
These data are made available on
MAST\footnote{\url{https://archive.stsci.edu/kepler/}} and, for each target,
we downloaded the full set of light curve files provided by Data Release 24
\citep{Thompson:2015}.
From these files, we extracted the PDC time series and split them into
``sections'' with no more than two contiguous missing or flagged data points.
The PDC light curves have been corrected for the instrumental effects caused
by the spacecraft using a data-driven model of the focal plane
\citep{Stumpe:2012, Smith:2012}.
Crucially, an attempt is also made by the PDC procedure to remove sharp
instrumental artifacts like ``sudden pixel sensitivity dropouts (SPSDs)''.

@@

\paragraph{Terminology}

Any remaining missing data points are filled in using linear interpolation
resulting in sections of contiguous time series measurements with nearly
uniform time sampling.
This procedure yields about 150 \emph{light curve sections} for the typical
\kepler\ target.

@@

We use the \project{scikit-learn}\citep{Pedregosa:2011} implementation of the
RF classification algorithm\footnote{Specifically we use the
\texttt{RandomForestClassifier} object;
\url{http://scikit-learn.org/stable/modules/generated/sklearn.ensemble.RandomForestClassifier.html}.}.
